---
title: "R Notebook"
output: html_notebook
---

```{r}
library(data.table)
library(tidyverse)
library(lubridate)
```

Read the data in

```{r}
cities <- fread('../data-raw/cities.csv')
head(cities)
```

```{r}
dim(cities)
```

Data about districts (part of cities)

```{r}
districts <- fread('../data-raw/districts.csv')
dim(districts)
head(districts)
```

Read the main data set that we can use for modelling purposes

```{r}
otodom <- fread("../data-raw/train-small1.csv")
str(otodom,1)
```

Check the summary of the data

```{r}
summary(otodom)
```

```{r}
head(otodom)
```

## outline

1. change created_at_first to date - time format and extract day and hour -- lubridate package
2. extracting relevant information from params variable -- stringr package (part of tidyverse), stringi (by Marek Gągolewski)


### working with dates

```{r}
otodom <- otodom %>% 
  mutate(created_at_first = ymd_hms(created_at_first),
         yq = quarter(created_at_first, with_year = T))

head(otodom)
```

See how many observations we have by quarters and by category

```{r}
otodom %>% 
  count(yq, category, is_business) %>%
  ggplot(data = ., aes(x = yq, y = n, color = category, group = category)) +
  geom_line() + 
  facet_wrap(~is_business)
```

Add information about the year and month

```{r}
otodom <- otodom %>% 
  mutate(year = year(created_at_first),
         month = month(created_at_first),
         year_month = round_date(created_at_first, unit = "month", week_start = 1))

otodom %>%
  select(id, created_at_first, year, month, year_month)
```

### working with strings

```{r}
head(otodom)
```

Locate the longest string in params

```{r}
longest_string <- which.max(str_length(otodom$params))
dummy_string <- otodom$params[longest_string]
dummy_string
```

```{r}
str_extract(string = dummy_string, pattern = "<br>[a-z_]+<=>")

str_extract_all(string = dummy_string, pattern = "<br>[a-z_]+<=>")
```

```{r}
str_split(string = dummy_string, pattern = "<br>")
```

Extract information about floor area and numer of rooms

```{r}
otodom %>%
  select(id, price, params) %>%
  mutate(floor_area = str_extract(params, "m<=>\\d{1,}(\\.\\d{1,2})?"),
         floor_area = str_extract(floor_area, "\\d{1,}(\\.\\d{1,2})?"),
         floor_area = as.numeric(floor_area),
         rooms_num = str_extract(params, "rooms_num<=>\\d{1,}"),
         rooms_num = str_extract(rooms_num, "\\d{1,}"),
         rooms_num = as.numeric(rooms_num)) %>%
  select(id, floor_area, rooms_num, price, params) -> otodom_small

otodom_small
```

```{r}
boxplot(log(price) ~ rooms_num, data = otodom_small)
```

```{r}
plot(log(otodom_small$floor_area), log(otodom_small$price))
```

```{r}
cor(log(otodom_small$floor_area), log(otodom_small$price), m = "s")
cor(log(otodom_small$floor_area), log(otodom_small$price), m = "p")
```

