---
title: "R Notebook"
output: html_notebook
---

```{r}
library(data.table) ## reading, processing, handling 
library(VIM) ## vis and imputation
library(validate) ## validation of rules 
library(simputation) ### user-friendly imputation
library(tidyverse)
library(yaImpute) ## verify after the lecture
```

Read the data

```{r}
otodom <- fread("../data-raw/train-small1.csv")
head(otodom)
```

We use function `aggr` to see the missing data within our data.frame

WORKS ONLY IF MISSING DATA IS CODED NA

If you have "NULL", "", -1, 0, -712i42

```{r}
miss_report <- aggr(x  = otodom)
```

```{r}
#otodom[,.N, keyby = city_id]
otodom %>% count(city_id)
```

Extract floor area and number of rooms

```{r}
otodom <- otodom %>%
  mutate(floor_area = str_extract(params, "m<=>\\d{1,}(\\.\\d{1,2})?"),
         floor_area = str_extract(floor_area, "\\d{1,}(\\.\\d{1,2})?"),
         floor_area = as.numeric(floor_area),
         rooms_num = str_extract(params, "rooms_num<=>\\d{1,}"),
         rooms_num = str_extract(rooms_num, "\\d{1,}"),
         rooms_num = as.numeric(rooms_num))

aggr(otodom %>% select(floor_area, rooms_num))

summary(otodom %>% select(floor_area, rooms_num))
```

- first: quality, remove erronous observations
- then imputation

Proceed with imputation of missing data in number of rooms with hot-deck method

hotdeck - function from VIM package

```{r}
otodom_imp <- hotdeck(data = otodom,
                      variable = "rooms_num",
                      domain_var = c("region_id", "city_id", "category"))
head(otodom_imp)
```

```{r}
otodom_imp %>% count(rooms_num_imp)
```

```{r}
otodom_imp %>%
  count(rooms_num_imp, rooms_num) %>%
  add_count(rooms_num_imp, wt = n, name = "total") %>%
  mutate(p = n/ total) %>%
  ggplot(data = ., aes(x = rooms_num, y = p, 
                       color = rooms_num_imp, fill = rooms_num_imp)) +
  geom_col(position = position_dodge())
```

Comment to this plot: we did it wrong! Mainly because of variables that we used, this is the problem. 

Imputation will work only if you use highly correlated variables with the one that you impute.

