---
title: "Quantile regression - introduction"
author: "Maciej Beręsewicz"
output: html_notebook
---

# Packages

```{r}
library(quantreg) ## install.packages("quantreg")
library(data.table)
library(tidyverse)
library(lubridate)
```


# Data 

```{r}
train <- fread("~/git/zbiory/otodom/train-small3.csv", 
               col.names = c("id", "created_at_first", "category",
                             "is_business", "district_id", "city_id", "region_id", 
                             "params", "title", "price")) %>%
  mutate(floor_area = str_extract(params, "m<=>\\d{1,}(\\.\\d{1,2})?"),
         floor_area = str_extract(floor_area, "\\d{1,}(\\.\\d{1,2})?"),
         floor_area = as.numeric(floor_area),
         rooms_num = str_extract(params, "rooms_num<=>\\d{1,}"),
         rooms_num = str_extract(rooms_num, "\\d{1,}"),
         rooms_num = as.numeric(rooms_num),created_at_first = ymd_hms(created_at_first),
         yq = quarter(created_at_first, with_year = T))

head(train)

```

# Quantile regression

## Literature

## Introduction

In quantile regression we are interested in modelling the $\tau$ quantile of target $Y$ variable instead of mean.

More formally, any real valued random variable  $Y$, may be characterized by its distribution function


$$
F(y) = Prob(Y \leq y)
$$

where for any $0 < \tau < 1$,

$$
Q(\tau) = \inf\{y : F(y) \geq \tau \}
$$

is called the $\tau$-th quantile of $Y$. The median, $Q(1/2)$, plays the central role. Like the distribution function, the quantile function provides a complete charicterization of the random variable $Y$.

The quantiles may be formulated as the solution to a simple optimization problem. For any $0 < \tau < 1$, define a picewise linear "check function"

$$
\rho_\tau(u) = u (\tau - I( u < 0))
$$

Minimizing the expectation of $\rho_\tau(Y-\xi)$ with respect to $\xi$ yelds solutions, $\hat{\xi}(\tau)$, the smallest of which is $Q(\tau)$ defined above.


The same analogue of $Q(\tau)$, based on a random sample, $\{y_1, y_2, ..., y_n \}$ of $Y$'s is called the $\tau$-th sample quantile, and may be found by solving

$$
\min\limits_{\xi in \boldsymbol{R}} \sum_{i=1}^n \rho_\tau(y_i - \xi).
$$

If we assume the linear conditional quantile function $Q_Y(\tau|X=x) = \boldsymbol{x}'_i\boldsymbol{\beta}$, we get

$$
\hat{\beta}(\tau) = \operatorname{arg\,min}\limits_{\boldsymbol{\beta} in \boldsymbol{R}^p} \sum_{i=1}^n\rho_{\tau}(y_i- \boldsymbol{x}'_i\boldsymbol{\beta}).
$$
## Quantile regression in R

The main package for quantile regression in R is `quantreg` which implements the following models

1. `quantreg::rq` -- linear quantile regression
2. `quantreg::nlrq` -- nonlinear quantile regression


## Example on real estate market


## linear regression

```{r}
head(train)
train <- train %>% mutate(region_id = as.factor(region_id),
                          city_id = as.factor(city_id))
```


```{r}
train <- train %>% filter(price > quantile(price,0.03), price < quantile(price,0.97))

model1 <- lm(formula = price ~ floor_area + rooms_num + is_business + category + region_id, data = train)
summary(model1)
round(coef(model1),4)
```

## quantile regression

We will start with the median

```{r}
model2 <- rq(formula = price ~ floor_area + rooms_num + is_business + category + region_id, tau = 0.5, data = train)
summary(model2)
```

```{r}
mean(sqrt(model1$residuals^2)) ## linear model
mean(sqrt(model2$residuals^2)) ## quantile model
```

Quantile regression for other percentiles

```{r}
model3 <- rq(formula = price ~ floor_area + rooms_num + is_business + category, tau = c(0.25, 0.5, 0.75), data = train)
summary(model3)
```

```{r}
plot(model3)
```

```{r}
model4 <- rq(formula = price ~ floor_area + rooms_num + is_business + category + region_id, 
             tau = seq(0.05,0.95,0.05), 
             data = train)
plot(model4)
```

Extract coefficients

```{r}
model4_result <- model4 %>% summary()

mode(model4_result)
length(model4_result)

coef(model4_result[[16]])

```

